---
title: "intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(transcribe)
```

# Introduction

This vignette introduces the **transcribe** package. The package provides an R interface to transcribe audio using Whisper, with optional post‐processing via Ollama.

## Installation

```{r, eval=FALSE}
# Install from GitHub (if not on CRAN yet)
# install.packages("remotes")
remotes::install_github("YourUserName/transcribe")
```

## Basic Usage

Below is a minimal example showing how to transcribe an audio file:

```{r, eval=FALSE}
# library(transcribe) # Uncomment once your package is installed
# transcript <- transcribe_audio(
#   input_path = "path/to/audio.wav",
#   language = "en",
#   whisper_model_name = "large-v3-turbo",
#   processed = TRUE,
#   ollama_model = "llama3.2"
# )
# cat(transcript)
```

## Plumber Integration

You can serve a web interface via **plumber**:

```{r, eval=FALSE}
# library(plumber)
# plumber::plumb("inst/plumber/api.R")$run()
```

Open your browser at `http://127.0.0.1:port` to see the interface.

## Conclusion

This vignette covered the basics of installing and using **transcribe** for audio transcription. See the README or additional vignettes for more details.
